### 2021-05-24 ms/ms.tex
Update manuscript to include Geir comments and prior sensitivity text

\usepackage[round,sectionbib]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}
\bibliographystyle{fishfishnourl}

@@

\textbf{Sebasti\'{a}n A. Pardo\textsuperscript{1*}, 
        Geir H. Bolstad\textsuperscript{2}, 
        J. Brian Dempson\textsuperscript{3}, 
        Julien April\textsuperscript{4}, 
        Ross A. Jones\textsuperscript{5}, 
        %Martha J. Robertson\textsuperscript{3}, 
        Dustin Raab\textsuperscript{6}, 
Jeffrey A. Hutchings\textsuperscript{1}}

\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}

@@

Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
Atlantic are primarily attributed to decreases in survival at sea. 
Whether these changes in marine survival are highly variable among populations
has not been straightforward to estimate because of
the need to model data on both migrating smolts and returning adults and to
simultaneously estimate multiple parameters. We fit a hierarchical Bayesian
maturity schedule model to data spanning between 19 and 42 years for seven populations in eastern 
Canada. We estimate numbers of out-migrating smolts, 
survival in the first and second year at sea, and the proportion returning after one year. 
Trends in survival at sea were not consistent among populations, and we observe both
positive, negative and no correlations in the trends. This suggest that large
scale patterns of changes in marine survival are not necessarily representative for
individual populations. 
Variation in return abundances was mostly explained by marine survival in the
first winter at sea in all but one population.
However, variation in the other components were not negligible and
their relative importance differed among populations. Variation in number of
out-migrating smolts had a moderate contribution on returns and probability of returning after one year was
important in some populations. 
% Trends were not consistent among rivers. 
% Since 1990, \So increased at Western Arm Brook (NL)
% and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
% de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
% declined at LaHave River (NS), and fluctuated at Nashwaak
% River (NB). Estimates of \St were highly uncertain, particularly for
% 1SW-dominated populations; \Pg was generally stable. 
% These results challenge the narrative that marine survival has changed in a
% temporally consistent manner among spatially disparate populations. 
If salmon populations do not respond in a consistent, uniform manner to
changing environmental (including oceanic) conditions throughout their range,
future research initiatives should explore why.

@@

Given the logistical challenges associated with estimating at-sea survival for
individual populations, it is not surprising that the number of studies that
have estimated temporal trends has been limited. 
The evidence for widespread declines in marine survival mostly comes from studies
over broad spatial scales that do not include empirical smolt abundance data, and
rely on highly variable stock-recruitment relationships.
An additional limitation has been the derivation of
proxies (e.g., return rates), rather than direct model-based estimates, of
marine survival.
\citet{Chaput2012a}, for example, examined the return rate of smolts to adult salmon 
as a metric of marine survival, finding that most Canadian populations 
had experienced declining return rates. 
However, with the exception of one-sea-winter (1SW) dominated populations 
(such as most populations in Newfoundland) where return rates 
closely approximate marine survival, return rates cannot directly be interpreted as marine survival rates and 
examination of trends in return rates alone can mask changes in
differential survival during different years at sea, as well as changes in the
proportion of adults returning after one or two years at sea \citep{Hubley2011}.

@@

\begin{equation}
    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,y,s})}}{\sum_{s}{N_{r,t,s}}}
\end{equation}

where $p_{r,t,a}$ is the proportion of annual returns in river $r$, year $t$,
and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,t,s,a}$ is the
number of samples in river $r$, year $t$, of returning age $a$, and of size
group $s$; $n_{r,t,s}$ is the total number of samples in river $r$, year $t$,
and of size group $s$; and $N_{r,t,s}$ is the returns of salmon
in river $r$, year $t$, and of size group $s$.
Ideally, we would account the uncertainty in the conversion between
size group and returning age in the model, but given limitations of the Bayesian 
sampler we used, we incorporated this uncertainty indirectly as an estimate
of the variance of annual log-smolt abundance.

@@

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate annual
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for the two marine survival parameters and the maturation
parameter in a Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter (2SW) adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg).

Our model does not include repeat spawners and assumes that no fish spend
three or more winters at sea before returning to spawn for the first time.
The model also assumes that mortality in the second winter at sea (\St) for 2SW returns
is additive to mortality in the first winter at sea in the previous year, 
and therefore does not account for differences in environmental conditions experienced
between 1SW and 2SW fish of the same smolt cohort during their overlapping first year at sea.
In other words, our model assumes that the decision of returning occurs just before
actually being counted as returns and that \St is additional mortality in
the subsequent year.

@@

\begin{equation}
log(smolts_{obs,t,r}) = log(smolts_{true,t,r}) + \epsilon_{t,r},
\end{equation}

where $smolts_{true,t,r}$ are the true smolt abundances for year $t$ and river
$r$, and $\epsilon_{t,r}$ is the error term, which is calculated from the yearly 
coefficient of variation in the empirically derived
smolt estimates (see Table S1 in the Supplementary material). 
Where available, we used population-specific measurement error estimates for smolt abundances; if not 
available, we set measurement error between 5 and 10\% (see Table S2 in the Supplementary material). 
The log-transformed true smolt abundances are
normally distributed around a population-level mean and standard deviation:

\begin{equation}
log(smolts_{true,t,r}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
\end{equation}

@@

\begin{align}
    R_{1,t} &= smolts_{true,t-1} * S_{1,t} * P_{g,t} \label{eq:1}, \\
    R_{2,t+1} &= smolts_{true,t-1} * S_{1,t} * (1 - P_{g,t}) * S_{2,t+1}, \label{eq:2}
\end{align}

where $R_{1,t}$ and $R_{2,t+1}$ are the estimated abundances of 1SW and 2SW
salmon returning in years $t$ and $t+1$, respectively, $smolts_{true,t-1}$ is the
estimated number of out-migrating smolts in year $t-1$, $S_{1,t}$ is the proportion of
salmon surviving in their first year ($t$) at sea, $P_{g,t}$ is the proportion of
salmon that return to spawn at year $t$, $S_{2,t+1}$ is the survival in their
second year at sea of the same cohort of salmon who did not return to spawn at
year $t$.

However, to allow for normally and log-normally-distributed parameters we
log-transform equations~\ref{eq:1} and~\ref{eq:2} to obtain

\begin{align}
    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(Pr_{r,t}) - Z_{1,r,t} \label{eq:3}, \\
    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - Pr_{r,t})  - Z_{2,r,t+1} \label{eq:4} 
\end{align}

@@

\begin{align}
log(R_{obs,1,t,r}) &\sim Normal(log(R_{1,t,r}), \epsilon_{1,t,r}), \\
log(R_{obs,2,t,r}) &\sim Normal(log(R_{2,t,r}), \epsilon_{2,t,r}) \label{eq:5} 
\end{align}

where $R_{obs,1,t,r}$ and $R_{obs,2,t,r}$ are the observed return estimates
for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,t,r}$
and $\epsilon_{2,t,r}$ are the process error terms. 
These error terms are estimated indirectly by approximating the abundance of 1SW and 2SW returns
from hypergeometric distributions based on the scale sample data. 
We were able to obtain scale data for all but two of the populations (LaHave and Nashwaak).
See the Supplementary Materials for a description of the methodology and the resulting estimates.

@@

\begin{align}
 S_{1} &= e^{-Z_1}, \\
 S_{2} &= e^{-Z_2}. 
\end{align}

@@

\begin{align}
Z_1 &\sim logNormal(1, 0.22),   \\ 
Z_2 &\sim logNormal(0.2, 0.3).
\end{align}

These priors, when converted to yearly survival, are roughly constrained between
0 and 0.2 for \So and between 0.2 and 0.5 for \St (Fig.~\ref{fig:priors}a). They are
informative only to the extent that they limit the estimates of marine survival
to what we considered to be biologically realistic.

We estimate population-level mean logit \Pg values around which the yearly \Pg
values are normally distributed. We specify different informative hyperpriors
for \prmu and \prsig based on whether the population is 1SW-dominated (i.e.
with a proportion of 1SW fish in the total returns greater than 0.9) or not:

\begin{align}
    logit(P_{g,t}) &\sim Normal(\mu_{Pg,r}, \sigma_{Pg,r}) \\
    \mu_{Pg,r} &\sim 
    \begin{cases}
       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
   \end{cases} \\
    \sigma_{Pg,r} &\sim halfNormal(0, 1).
\end{align}

The three Newfoundland populations (WAB, Campbellton and Conne) are 
1SW-dominated, while the other four have the more generic priors for $logit(P_{g,t})$.
These priors, when converted back to a proportion, are narrowly constrained
for 1SW-dominated populations, but relatively wide for all other populations
(Fig.~\ref{fig:priors}b).

@@

The model was written in Stan \citep{Carpenter2017} and run in R version 3.6.1
\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
\citep{StanDevelopmentTeam2019}.
The model was run with three chains and 3,000 iterations, with the first 1,500
discarded as a burn-in. The models were considered to have converged when the
$\hat R$ of all parameters were lower than 1.03 and the effective sample size 
were higher than 500.

@@

Lastly, to determine what parameters best explains the number of returns in
the model (including the model uncertainty), we estimated a pseudo-$R^2$ by
calculating the squared correlation between the estimated parameters and the
estimated 1SW and 2SW returns ($R_{1,r,t}$ and $R_{2,r,t+1}$ in
equation~\ref{eq:4}).

@@

Long-term trends among populations also varied: Campbellton,
Saint-Jean, and Western Arm Brook populations showed increases in \So
over time, Trinit\'{e}, Conne, and LaHave populations showed decreases,
while at Nashwaak there was an increase in median \So during the early
2000s but a decrease since 2010. Annual estimates of \So had little
uncertainty for one sea-winter dominated populations (Conne, Campbellton, and
WAB), with the exception of the 2004-2011 periord in the Campbellton where
there were very few scale samples, and hence high process error; but in general, 
uncertainty was higher (i.e. wider credible intervals) in the other
populations.

@@

Estimates of \Pg were mostly stable across time, except for the LaHave and
Nashwaak populations; the estimates of \Pg were slightly lower in the last
four years than in the previous ones, while in the Nashwaak the posterior
estimates of \Pg in 2012 were much lower than in all other years
(Fig.~\ref{fig:s2-faceted}). Variation in yearly estimates was highest in the
LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
populations.

@@

Estimates of the population-level mean (\prmu) and standard deviation (\prsig) of proportion returning after one year at sea varied considerably among populations. 
For all
three 1SW-dominated populations (Campbellton, Conne, and WAB), estimates of \prmu
were very close to 1.0 and had little variation in \prsig (Fig~\ref{fig:prmu-post}; Table S7 in the Supplementary Materials).
Estimates of \prmu were the lowest for the two QC populations, particularly the Saint-Jean (median \prmu = 0.11).
Estimates of \prmu for the Nashwaak and the LaHave populations were close to 0.5, with these two populations having 
the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post};  Table S8 in the Supplementary Materials).

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=1.0\linewidth]{figures/logitpr-mu-sigma-post.png}
    \caption{Posterior probability distributions resulting from the population-level estimates of the proportion returning after one winter at sea based on
        the parameters \prmu and \prsig. Gray lines denote population-specific median estimates of \prmu and \prsig, while
       the colored lines represent a sample of 20 posterior iteration draws for each population.} 
   \label{fig:prmu-post} 
\end{figure}

@@

Roughly half (11) out of the 21 pair-wise correlations were significantt (p-value below 0.05)
(Fig.~\ref{fig:s1-corr}). When looking at the direction of the correlations,
these spanned both positive and negative coefficients. All seven populations had both
positive and negative correlations among them.
The corrected pair-wise correlations in $Z_1$ were almost identical to the
uncorrected ones as the error variance and covariance terms were very small.
The correlations between populations in the region were positive between Nashwaak and LaHave (Scotia-Fundy region), non-significant between the Trinit\'{e} and Saint-Jean (QC), 
non-significant between WAB and Campbellton but negative between Conne and WAB (NL).
The Campbellton had mostly non-significant pairwise correlations with other populations, even though there are both positive and negative correlation coefficients, except with the Saint-Jean
where the correlation between $Z_1$ trends is positive and significant.
Note that the uncertainty in these correlations reflect the uncertainty in the
measurements, but not in the process (in this respect they are equivalent to
sample variance and not population variance).

@@

As is expected from Murphy's method, there is a strong negative correlation
between survival in the second year at sea (i.e. $-Z_2$) and proportion returning
as 1SW ($log(P_g)$, Fig.~\ref{fig:cor-params}). Some populations (Campbellton,
LaHave, Nashwaak, and Saint-Jean) show a negative correlation between the
estimated number of smolts and survival in the first year at sea, while the
others fluctuate around zero. There is overall a positive correlation between
survival in the first year (i.e. $-Z_1$) and proportion returning as 1SW,
except for the Campbellton population, where this correlation is negative.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/corrected-cor-ALL.png}
    \caption{Correlations among estimated model parameters. \emph{log(smolts)} = estimated smolt abundances in log-space, $log(P_g)$ = logit-transformed probability of
        returning after one year at sea, $-Z_1$ = marine survival in the first year at sea, and $-Z_2$ = marine survival in the second year at sea.
        Note the correlations derived from negative instantaneous total mortalities $-Z_1$ and $-Z_2$ are representative to those of marine survival $S_1$ and $S_2$.}
    \label{fig:cor-params} 
\end{figure}

>>>>

\usepackage[round,sectionbib]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}
\bibliographystyle{icesjms}

@@

\textbf{Sebasti\'{a}n A. Pardo\textsuperscript{1*},
        Geir H. Bolstad\textsuperscript{2},
        J. Brian Dempson\textsuperscript{3},
        Julien April\textsuperscript{4},
        Ross A. Jones\textsuperscript{5},
        %Martha J. Robertson\textsuperscript{3},
        Dustin Raab\textsuperscript{6},
Jeffrey A. Hutchings\textsuperscript{1,7,8}}

\noindent\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{7} Institute of Marine Research, Flødevigen Marine Research Station, N-4817 His, Norway\\}
\small{\textsuperscript{8} Centre for Coastal Research, University of Agder, N-4604 Kristiansand, Norway\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}

@@

Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
Atlantic are primarily attributed to decreases in survival at sea.
However, comparing trends in marine survival among populations
is challenging as data on both migrating smolts and returning adults is sparse and
models are difficult to parameterize due to their varied life histories. We fit a hierarchical Bayesian
maturity schedule model to data from seven populations in eastern
Canada to estimate numbers of out-migrating smolts,
survival in the first and second year at sea, and the proportion returning after one year.
Trends in survival at sea were not consistent among populations; we observe
positive, negative and no correlations in these, suggesting that large
scale patterns of changes in marine survival are not necessarily representative for
individual populations.
Variation in return abundances was mostly explained by marine survival in the
first winter at sea in all but one population.
However, variation in the other components were not negligible and
their relative importance differed among populations.
If salmon populations do not respond in a uniform manner to
changing environmental conditions throughout their range,
future research initiatives should explore why.

% Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
% Atlantic are primarily attributed to decreases in survival at sea.
% Whether these changes in marine survival are highly variable among populations
% has not been straightforward to estimate because of
% the need to model data on both migrating smolts and returning adults and to
% simultaneously estimate multiple parameters. We fit a hierarchical Bayesian
% maturity schedule model to data spanning between 19 and 42 years for seven populations in eastern
% Canada. We estimate numbers of out-migrating smolts,
% survival in the first and second year at sea, and the proportion returning after one year.
% Trends in survival at sea were not consistent among populations, and we observe both
% positive, negative and no correlations in the trends. This suggest that large
% scale patterns of changes in marine survival are not necessarily representative for
% individual populations.
% Variation in return abundances was mostly explained by marine survival in the
% first winter at sea in all but one population.
% However, variation in the other components were not negligible and
% their relative importance differed among populations. Variation in number of
% out-migrating smolts had a moderate contribution on returns and probability of returning after one year was
% important in some populations.
% % Trends were not consistent among rivers.
% % Since 1990, \So increased at Western Arm Brook (NL)
% % and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
% % de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
% % declined at LaHave River (NS), and fluctuated at Nashwaak
% % River (NB). Estimates of \St were highly uncertain, particularly for
% % 1SW-dominated populations; \Pg was generally stable.
% % These results challenge the narrative that marine survival has changed in a
% % temporally consistent manner among spatially disparate populations.
% If salmon populations do not respond in a consistent, uniform manner to
% changing environmental (including oceanic) conditions throughout their range,
% future research initiatives should explore why.

@@

Given the logistical challenges associated with estimating at-sea survival for
individual populations, it is not surprising that the number of studies that
have estimated temporal trends has been limited.
The evidence for widespread declines in marine survival mostly comes from studies
over broad spatial scales that do not include empirical smolt abundance data, and
rely on highly variable stock-recruitment relationships.
An additional limitation has been the derivation of
proxies (e.g., return rates), rather than direct model-based estimates, of
marine survival.
For example, \citet{Chaput2012a} examined return rates (i.e., ratio of adult salmon to outgoing smolts)
as a metric of marine survival, finding that most Canadian populations
had experienced declining return rates.
However, with the exception of one-sea-winter (1SW) dominated populations
(such as most populations in Newfoundland) where return rates
closely approximate marine survival, return rates cannot directly be interpreted as marine survival rates and
examination of trends in return rates alone can mask changes in
differential survival during different years at sea, as well as changes in the
proportion of adults returning after one or two years at sea \citep{Hubley2011}.

@@

\begin{equation}
    p_{r,t,a} = \frac{\sum_{s}{(\frac{n_{r,t,s,a}}{n_{r,t,s}} * N_{r,t,s})}}{\sum_{s}{N_{r,t,s}}}
\end{equation}

where $p_{r,t,a}$ is the proportion of annual returns in river $r$, year $t$,
and of spawning history $a$ (either 1SW or 2SW returns); $n_{r,t,s,a}$ is the
number of samples in river $r$, year $t$, of returning age $a$, and of size
group $s$; $n_{r,t,s}$ is the total number of samples in river $r$, year $t$,
and of size group $s$; and $N_{r,t,s}$ is the returns of salmon
in river $r$, year $t$, and of size group $s$.
For years where scale data of a given size group is lacking,
we averaged the proportions of annual returns
for the closest ten years for which there was data.
We incorporated the uncertainty in the conversion between
size group and returning age in the model indirectly as an estimate
of the variance of annual log-1SW and 2SW return abundance (see Supplementary Materials).

@@

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate annual
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome by
using informative priors for the two marine survival parameters and the maturation
parameter in a Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter (2SW) adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg).

Our model does not include repeat spawners and assumes that no fish spend
three or more winters at sea before returning to spawn for the first time.
Assuming that there are no age 3+ maiden spawners is likely to result in
slightly lower S1 and S2 estimates as these older maiden spawners that would
be returning in later years are assumed to have died. Nonetheless, in the
populations examined these fish comprise a very small fraction of maiden spawners,
and we do not expect these to substantially affect our results.
The model also assumes that mortality in the second winter at sea (\St) for 2SW returns
is additive to mortality in the first winter at sea in the previous year,
and therefore does not account for differences in environmental conditions experienced
between 1SW and 2SW fish of the same smolt cohort during their overlapping first year at sea.
In other words, our model assumes that the decision of returning occurs just before
actually being counted as returns and that \St is additional mortality in
the subsequent year.

@@

\begin{equation}
log(smolts_{obs,r,t}) = log(smolts_{true,r,t}) + \epsilon_{r,t},
\end{equation}

where $smolts_{true,r,t}$ are the true smolt abundances for river
$r$ and year $t$, and $\epsilon_{r,t}$ is the error term, which is calculated from the yearly
coefficient of variation in the empirically derived
smolt estimates (see Tables S4-6 in the Supplementary Material).
Where available, we used population-specific measurement error estimates for smolt abundances; if not
available, we set the measurement error at 5 or 10\%, depending on whether the smolt
abundance estimates originated from direct counts or mark-recapture studies, respectively
(see Table S5 in the Supplementary Material).
The log-transformed true smolt abundances are
normally distributed around a population-level mean and standard deviation:

\begin{equation}
log(smolts_{true,r,t}) \sim Normal(\mu_{smolts,r}, \sigma_{smolts,r})
\end{equation}

@@

\begin{align}
    R_{r,1,t} &= smolts_{true,r,t-1} * S_{1,r,t} * P_{g,r,t} \label{eq:1}, \\
    R_{r,2,t+1} &= smolts_{true,r,t-1} * S_{1,r,t} * (1 - P_{g,r,t}) * S_{2,r,t+1}, \label{eq:2}
\end{align}

where $R_{1,r,t}$ and $R_{2,r,t+1}$ are the estimated abundances of 1SW and 2SW
salmon returning for river $r$ in years $t$ and $t+1$, respectively, $smolts_{true,r,t-1}$ is the
estimated number of out-migrating smolts for river $r$ in year $t-1$, $S_{1,r,t}$ is the proportion of
salmon surviving in their first year ($t$) at sea for river $r$, $P_{g,t}$ is the proportion of
salmon that return to spawn at year $t$ for river $r$, and $S_{2,r,t+1}$ is the survival in their
second year at sea of the same cohort of salmon who did not return to spawn at
year $t$ for river $r$.

We log-transform equations~\ref{eq:1} and~\ref{eq:2} so the model
is linear on the log-scale:

\begin{align}
    log(R_{1,r,t}) &= log(smolts_{r,t-1}) + log(P_{g,r,t}) - Z_{1,r,t} \label{eq:3}, \\
    log(R_{2,r,t+1}) &= log(smolts_{r,t-1}) - Z_{1,r,t} + log(1 - P_{g,r,t})  - Z_{2,r,t+1} \label{eq:4}
\end{align}

@@

\begin{align}
log(R_{obs,1,r,t}) &\sim Normal(log(R_{1,r,t}), \epsilon_{1,r,t}), \\
log(R_{obs,2,r,t}) &\sim Normal(log(R_{2,r,t}), \epsilon_{2,r,t}) \label{eq:5}
\end{align}

where $R_{obs,1,r,t}$ and $R_{obs,2,r,t}$ are the observed return estimates
for year $t$ and river $r$ of 1SW and 2SW fish, respectively, $\epsilon_{1,r,t}$
and $\epsilon_{2,r,t}$ are the process error terms.
These error terms are estimated by approximating the abundance of 1SW and 2SW returns
from hypergeometric distributions based on the scale sample data.
We assume that the error in the total number of returns is zero;
the error in the total number of returns is likely minor
compared to the error in the number of outmigrating smolts and the error due to
a small scale sample size. Hence, ignoring the error in the returns is
unlikely to change our results significantly.
We were able to obtain scale data for all but two of the populations (LaHave and Nashwaak), and we
bootstrapped a distribution of abundances of annual 1SW and 2SW returns for each populations from hypergeometric distributions,
from which we estimated the variance in log-space.
See the Supplementary Materials for a description of the methodology and the resulting estimates.

@@

\begin{align}
 S_{1,r,t} &= e^{-Z_{1,r,t}}, \\
 S_{2,r,t} &= e^{-Z_{2,r,t}}.
\end{align}

@@

\begin{align}
Z_{1,r,t} &\sim logNormal(1, 0.22),   \\
Z_{2,r,t} &\sim logNormal(0.2, 0.3).
\end{align}

These priors, when converted to yearly survival, are roughly constrained between
0 and 0.2 for \So and between 0.2 and 0.5 for \St (Fig.~\ref{fig:priors}a). They are
informative only to the extent that they limit the estimates of marine survival
to what we considered to be biologically realistic based on previous studies on
return rates of 1SW-dominated populations in NL, as well as the couple studies
that have explored survival in the second winter at sea of Atlantic salmon in
eastern Canada.
Although two published values are higher than 0.8 \citep{Chaput2003b,Hubley2011}, most estimates of S2 encompass a range between 0.1 and 0.6; this is the empirical basis for our prior.
While there is circularity in setting priors based on data for the same rivers,
and is far from ideal, these are the only data we have available to set priors
and the models used previously are considerably different to the one in the present study.

We estimate population-level mean probit \Pg values around which the yearly \Pg
values are normally distributed. We specify different informative hyperpriors
for \prmu and \prsig based on whether the population is 1SW-dominated (i.e.
with a proportion of 1SW fish in the total returns greater than 0.9) or not:

\begin{align}
    probit(P_{g,t}) &\sim Normal(\mu_{Pg,r}, \sigma_{Pg,r}) \\
    \mu_{P_{g,r}} &\sim
    \begin{cases}
       Normal(2.3, 0.4),  &\text{for 1SW-dominated populations} \\
       Normal(0, 2.8), &\text{for non-1SW-dominated populations} \\
   \end{cases} \\
    \sigma_{Pg,r} &\sim halfNormal(0, 1).
\end{align}

The three Newfoundland populations (WAB, Campbellton and Conne) are
1SW-dominated, while the other four have the more generic priors for $probit(P_{g,t})$.
These priors, when converted back to a proportion, are narrowly constrained
for 1SW-dominated populations, but relatively wide for all other populations
(Fig.~\ref{fig:priors}b).

@@

@@

Lastly, to determine what parameters best explains the number of returns in
the model (including the model uncertainty), we estimated the variance in returns explained by each parameter by
calculating the squared correlation between the estimated parameters and the
estimated 1SW and 2SW returns ($R_{1,r,t}$ and $R_{2,r,t+1}$ in
equations~\ref{eq:3} and~\ref{eq:4}).

The model was written in Stan \citep{Carpenter2017} and all analyses were run in R version 3.6.1
\citep{RCoreTeam2019} using the \texttt{rstan} package version 2.19.2
\citep{StanDevelopmentTeam2019}.
The model was run with three chains and 3,000 iterations, with the first 1,500
discarded as a burn-in. The models were considered to have converged when the
$\hat R$ of all parameters were lower than 1.03 and the effective sample size
were higher than 500. The data and code are available at \url{https://github.com/sebpardo/salmon-marine-survival}.

@@

Visual inspection of long-term trends among populations also varied: Campbellton,
Saint-Jean, and Western Arm Brook populations showed increases in \So
over time, Trinit\'{e}, Conne, and LaHave populations showed decreases,
while at Nashwaak there was an increase in median \So during the early
2000s but a decrease since 2010. Annual estimates of \So had little
uncertainty for one sea-winter dominated populations (Conne, Campbellton, and
WAB), with the exception of the 2004-2011 periord in the Campbellton where
there were very few scale samples, and hence high process error; but in general,
uncertainty was higher (i.e. wider credible intervals) in the other
populations.

@@

Estimates of \Pg were mostly stable across time, except for the LaHave and
Nashwaak populations; for the La Have the estimates of \Pg were slightly lower in the last
four years than in the previous ones, while in the Nashwaak the posterior
estimates of \Pg in 2012 were much lower than in all other years
(Fig.~\ref{fig:pr-faceted}). Variation in yearly estimates was highest in the
LaHave, Nashwaak, and Trinit\'{e}, and lowest in the 1SW-dominated
populations.

@@

Estimates of the population-level mean (\prmu) and standard deviation (\prsig)
of proportion returning after one year at sea varied considerably among populations.
For all
three 1SW-dominated populations (Campbellton, Conne, and WAB), estimates of \prmu
were very close to 1.0 and had little variation in \prsig (Fig~\ref{fig:prmu-post}; Table S10 in the Supplementary Materials).
Estimates of \prmu were the lowest for the two QC populations, particularly the Saint-Jean (median \prmu = 0.11).
Estimates of \prmu for the Nashwaak and the LaHave populations were close to 0.5, with these two populations having
the highest estimated values of \prsig, particularly the Nashwaak (Fig~\ref{fig:prmu-post};  Table S11 in the Supplementary Materials).

Uniquely for the Saint-Jean population, yearly variation in \St is being
estimated by the model but there is almost no variation in \Pg estimates
(Fig.~\ref{fig:pr-faceted}) and a very low estimate of \prsig (Table S11
in the Supplementary Materials). In other words, the variation in the relative
proportion of 1SW and 2SW returns for Saint-Jean is being manifested as
variation in \So and \St but not in \Pg rather than being manifested in \So
and \Pg but not \St, which is the case for the other six populations. This
difference is likely a result of the identifiability issue with the maturity
schedule method, which for the Saint-Jean population data, combined with the priors
used, results in the model estimating variation in \Pg rather than \St.

\ref{fig:pr-faceted}

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=1.0\linewidth]{figures/logitpr-mu-sigma-post.png}
    \caption{Posterior probability distributions resulting from the population-level estimates of the proportion returning after one winter at sea based on
        the parameters \prmu and \prsig. Gray lines denote population-specific median estimates of $P_{g,r}$, while
       the colored lines represent a sample of 20 posterior iteration draws for each population.}
   \label{fig:prmu-post}
\end{figure}

@@

Roughly half (11) out of the 21 pair-wise correlations were significant (p-value below 0.05)
(Fig.~\ref{fig:s1-corr}). When looking at the direction of the correlations,
these spanned both positive and negative coefficients. All seven populations had both
positive and negative correlations among them.
The corrected pair-wise correlations in $Z_1$ were almost identical to the
uncorrected ones as the error variance and covariance terms were very small.
The correlations between populations in the region were positive between Nashwaak and LaHave (Scotia-Fundy region), non-significant between the Trinit\'{e} and Saint-Jean (QC),
non-significant between WAB and Campbellton but negative between Conne and WAB (NL).
The Campbellton had mostly non-significant pairwise correlations with other populations, even though there are both positive and negative correlation coefficients, except with the Saint-Jean
where the correlation between $Z_1$ trends is positive and significant.
Note that the uncertainty in these correlations reflect the uncertainty in the
measurements, but not in the process (in this respect they are equivalent to
population variance and not sample variance).
Our choice of priors affected these correlations:
using weaker priors for \So and \St resulted in slightly different coefficients and significance thresholds (see Fig. S7 of Supplementary materials).

@@

As is expected from Murphy's method, there are strong negative, yet highly uncertain, correlations
between survival in the second year at sea (i.e. $-Z_2$) and proportion returning
as 1SW ($log(P_g)$, Fig.~\ref{fig:cor-params}). Some populations (Campbellton,
LaHave, Nashwaak, and Saint-Jean) show a negative correlation between the
estimated number of smolts and survival in the first year at sea, while the
others fluctuate around zero. There is overall a positive correlation between
survival in the first year (i.e. $-Z_1$) and proportion returning as 1SW,
except for the Campbellton population, where this correlation is negative.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/corrected-cor-ALL.png}
    \caption{Correlations among estimated model parameters. \emph{log(smolts)} = estimated smolt abundances in log-space, $probit(P_g)$ = probit-transformed probability of
        returning after one year at sea, $-Z_1$ = marine survival in the first year at sea, and $-Z_2$ = marine survival in the second year at sea.
        Note the correlations derived from negative instantaneous total mortalities $-Z_1$ and $-Z_2$ are representative to those of marine survival $S_1$ and $S_2$.}
    \label{fig:cor-params}
\end{figure}
